---
title: "Simulation Challenge"
subtitle: "Generative Models and Monte Carlo Simulation"
format:
  html: default

execute:
  echo: true
  eval: false
knitr:
  opts_chunk:
    python.reticulate: false
---

# 🎲 Simulation Challenge - Monte Carlo Analysis

## Challenge Overview

**Your Mission:** Create a comprehensive Quarto document that simulates one or two investment strategies, analyzes the results, and demonstrates your ability to present counter-intuitive findings compellingly. Then render the document to HTML and deploy it via GitHub Pages from a new repository called "simulationChallenge."

::: {.callout-warning}
## ⚠️ AI Partnership Required

This challenge pushes boundaries intentionally. You'll tackle problems that normally require weeks of study, but with Cursor AI as your partner (and your brain keeping it honest), you can accomplish more than you thought possible.

**The new reality:** The four stages of competence are Ignorance → Awareness → Learning → Mastery. AI lets us produce Mastery-level work while operating primarily in the Awareness stage. I focus on awareness training, you leverage AI for execution, and together we create outputs that used to require years of dedicated study.
:::

## The Investment Game 🎯

### Original Game Strategy

::: {#exm-ErgodicityEconomicsExample}
Imagine you are offered the following game and given a $1,000 budget in a special account to play the game: I will flip a coin, and if it comes up heads, we increase your account's balance by 50%; if it comes up tails, we reduce your account's balance by 40%. We are not only doing this once, but we will do it once per year until you turn 55. When you turn 55, you will receive the balance in your account.
:::

### Generative DAG Model for the Investment Game

```{python}
#| eval: false
#| echo: false
#| include: false
import daft

# Create the DAG directly without custom class
investmentDAG = daft.PGM(dpi=150, alternate_style="outer")

# Initial wealth (deterministic)
investmentDAG.add_node("W0", "Initial Wealth\n$W_0 = 1000$", x=4, y=2.5, aspect=5.4, 
                      alternate=True, plot_params={'facecolor': 'aliceblue'})

# Time t nodes  
investmentDAG.add_node("Ct", "Coin Flip t\n$C_t \\sim \\text{Bernoulli}(0.5)$", x=0, y=1, aspect=4, scale = 1.6,
                      plot_params={'facecolor': 'aliceblue'})
investmentDAG.add_node("Wt", "\nWealth t\n$W_t = 1.5 \\times W_{t-1}$ if $C_t = 1$\n$W_t = 0.6 \\times W_{t-1}$ if $C_t = 0$\n", x=4, y=1, scale = 1.6, aspect=4,
                      alternate=True, plot_params={'facecolor': 'aliceblue'})

# Add edges showing the relationships
investmentDAG.add_edge("W0", "Wt")
investmentDAG.add_edge("Ct", "Wt")

# Add plate around time-dependent nodes
investmentDAG.add_plate([-2, 0.5, 7.8, 1.0], label="t = 1, ..., N", shift=-0.1)
```

```{python}
#| eval: false
#| label: fig-investment-dag
#| fig-cap: Generative DAG model for the investment game showing how wealth evolves over time through coin flips
#| echo: false
investmentDAG.show()
```


## Challenge Requirements 📋

### Minimum Requirements for Any Points on Challenge

1. **Create a Quarto Document:** Write a concise quarto markdown file that includes a narrative of what you are doing along with the requested code, results, and visualizations of your simulations.

   ::: {.callout-tip}
   ## 💡 Pro Tip: Source File Reference
   
   **Need help with Quarto syntax or DAFT diagrams?** You can find the complete source `.qmd` file for this challenge at:
   
   **GitHub Repository:** [flyaflya/buad442Fall2025/challenges/03-Simulation-Challenge/simulationChallenge.qmd](https://github.com/flyaflya/buad442Fall2025/blob/main/challenges/03-Simulation-Challenge/simulationChallenge.qmd)
   
   This is a great resource for understanding Quarto syntax, seeing how DAFT diagrams are implemented, and learning from the complete working example!
   :::

2. **Render to HTML:** You must render the quarto markdown file to HTML.

3. **GitHub Repository:** The rendered HTML must be uploaded to a new GitHub repository called "simulationChallenge" in your Github account.

4. **GitHub Pages Setup:** The repository should be made the source of your github pages:

   - Go to your repository settings (click the "Settings" tab in your GitHub repository)
   - Scroll down to the "Pages" section in the left sidebar
   - Under "Source", select "Deploy from a branch"
   - Choose "main" branch and "/ (root)" folder
   - Click "Save"
   - Your site will be available at: `https://[your-username].github.io/simulationChallenge/`
   - **Note:** It may take a few minutes for the site to become available after enabling Pages

## Grading Rubric 🎓

### Questions to Answer for 75% Grade on Challenge

1. **Expected Value Analysis:** What is the "expected value" of your account balance after 1 coin flip for the original game?

2. **Expectation vs. Reality:** Is the expected value positive or negative? Do you expect your account to be worth more or less than $1,000 based on this result?

3. **Single Simulation:** Run one simulation showing the dynamics of your account balance over time. Make an object-oriented matplotlib OR ggplot2 plot showing your simulated account balance over time (i.e. as you age). Comment on the results, are you happy?

### Questions to Answer for 85% Grade on Challenge

4. **Multiple Simulations:** Run 100 simulations modelling the dynamics of your account balance over time. Make an object-oriented matplotlib OR ggplot2 plot showing a probability distribution of the 100 simulatedaccount balance at age 55. Comment on the results, are you happy? Why or why not?

### Questions to Answer for 95% Grade on Challenge

5. **Probability Analysis:** Based on the 100 simulations above, what is the probability that your account balance will be greater than $1,000 at age 55?

### Questions to Answer for 100% Grade on Challenge

6. **Strategy Comparison:** Run 100 simulations for the modified game strategy shown below in @exm-ErgodicityEconomicsExampleModified. What is the probability that your account balance will be greater than $10,000 at age 55? Is this probability higher or lower than the probability in the original game?

### Modified Game Strategy

::: {#exm-ErgodicityEconomicsExampleModified}
Imagine you are offered the following game and given a $1,000 budget in a special account to play the game: I will flip a coin, and if it comes up heads, we increase your bet by 50%; if it comes up tails, we reduce your bet by 40%. You must bet exactly 50% of your current account balance on each flip, and this 50% is locked in for each round. We are not only doing this once, but we will do it once per year until you turn 55. When you turn 55, you will receive the balance in your account.
:::

## Technical Implementation Preferences 💡

### Setting Up Your Analysis

**For R Users:**

- Use `tidyverse` for data manipulation
- Use `ggplot2` for visualizations
- Use `set.seed()` for reproducible results

**For Python Users:**

- Use `numpy` for numerical operations
- Use `pandas` for data manipulation
- Use `matplotlib` (object-oriented)
- Use `np.random.seed()` for reproducible results

### Visualization Preferences

- **Professional Styling:** Use consistent colors, clear labels, readable fonts, and informative titles

## Submission Checklist ✅

**Minimum Requirements (Required for Any Points):**

- [ ] Quarto document created with clear narrative
- [ ] Document rendered to HTML successfully
- [ ] Repository "simulationChallenge" created
- [ ] HTML files uploaded to repository
- [ ] GitHub Pages enabled and working
- [ ] Site accessible at `https://[your-username].github.io/simulationChallenge/`

**75% Grade Requirements:**

- [ ] Expected value calculations shown (Question 1)
- [ ] Expectation vs. reality analysis (Question 2)
- [ ] Single simulation with time series plot (Question 3)
- [ ] Clear interpretation of single simulation results

**85% Grade Requirements:**

- [ ] 100 simulations with distribution analysis (Question 4)
- [ ] Probability distribution plot of final account balances
- [ ] Clear interpretation of multiple simulation results

**95% Grade Requirements:**

- [ ] Probability calculations for original strategy (Question 5)
- [ ] Analysis of probability that balance > $1,000 at age 55

**100% Grade Requirements:**

- [ ] 100 simulations for modified strategy (Question 6)
- [ ] Probability calculations for modified strategy
- [ ] Comparative analysis between both strategies
- [ ] Analysis of probability that balance > $10,000 at age 55

**Code Quality (All Grades):**

- [ ] Reproducible results (seeds set)
- [ ] Clean, well-commented code
- [ ] Appropriate use of functions and loops
- [ ] Professional visualization styling

### Resources

- **Quarto Markdown:** [quarto.org/docs/authoring/markdown-basics.html](https://quarto.org/docs/authoring/markdown-basics.html)
- **Quarto Documentation:** [quarto.org/docs](https://quarto.org/docs)
- **R for Data Science:** [r4ds.had.co.nz](https://r4ds.had.co.nz)
- **Python Data Science Handbook:** [jakevdp.github.io/PythonDataScienceHandbook](https://jakevdp.github.io/PythonDataScienceHandbook)

### Getting Started Tips

::: {.callout-note}
## 🎯 Navy SEALs Motto

> "Slow is Smooth and Smooth is Fast"

*Take your time to understand the simulation mechanics, plan your approach carefully, and execute with precision. Rushing through this challenge will only lead to errors and confusion.*
:::

- **Browse [Essential Simulation Concepts](@sec-simulation-concepts):** This section will give you a good understanding of the concepts you need to know to complete the challenge.
- **Start Simple:** Begin with a single simulation to understand the mechanics
- **Document Everything:** Explain your reasoning and interpret your results
- **Forgetting to Set Seeds:** Always set random seeds for reproducible results
- **Total time to complete:** ~3-4 hours for the 100% grade ⏱️
- **Good luck, and remember simulation will steer you right even when intuition will steer you wrong!** 🎲

::: {.callout-warning}
## 💾 Important: Save Your Work Frequently!

**Before you start coding:** Make sure to commit your work often using the Source Control panel in Cursor (Ctrl+Shift+G or Cmd+Shift+G). This prevents the AI from overwriting your progress and ensures you don't lose your work.

**Commit after each major step:**

- After completing each simulation example
- After finishing each challenge question
- Before asking the AI for help with new code

**How to commit:**

1. Open Source Control panel (Ctrl+Shift+G)
2. Stage your changes (+ button)
3. Write a descriptive commit message
4. Click the checkmark to commit

*Remember: Frequent commits are your safety net!*
:::

## Essential Simulation Concepts 🎯 {#sec-simulation-concepts}

Before diving into the challenge, let's review the key simulation concepts you'll need. These examples will prepare you for the investment game analysis.

### 1. Simple Simulation: Coin Flip Game

Let's start with a basic coin flip simulation to understand the mechanics:

#### Generative DAG Model for the Simple Coin Flip Game

```{python}
#| eval: false
#| echo: false
#| include: false
import daft

# Create the DAG for simple coin flip game
simpleCoinDAG = daft.PGM(dpi=150, alternate_style="outer")

# Coin flip (stochastic node)
simpleCoinDAG.add_node("X", "Coin Flip\n$X \\sim \\text{Bernoulli}(0.5)$", x=1, y=1, aspect=4,
                      plot_params={'facecolor': 'aliceblue'})

# Winnings (deterministic node)
simpleCoinDAG.add_node("W", "Winnings\n$W = 100$ if $X = 1$ else $-100$", x=1, y=0, aspect=5.4,
                      alternate=True, plot_params={'facecolor': 'aliceblue'})

# Add edge showing the relationship
simpleCoinDAG.add_edge("X", "W")
```

```{python}
#| eval: false
#| label: fig-simple-coin-dag
#| fig-cap: Generative DAG model for the simple coin flip game showing the relationship between coin flip outcome and winnings
#| echo: false
simpleCoinDAG.show()
```

**Key Difference from Investment Game:** Unlike the investment game DAG (@fig-investment-dag) which models wealth evolution over multiple time periods with multiplicative changes, this simple coin flip DAG represents a single-period game with additive winnings. The investment game shows how wealth compounds over time ($W_t = 1.5 \times W_{t-1}$ or $W_t = 0.6 \times W_{t-1}$), while this simple game shows fixed winnings ($W = +100$ or $W = -100$) based on a single coin flip outcome.

::: {.panel-tabset}

### R

```{r}
#| label: simple-sim-r
#| fig-cap: R simulation of coin flip game
#| echo: true

suppressPackageStartupMessages(library(tidyverse))

# Set seed for reproducibility
set.seed(123)

# Number of simulations
n_sims <- 10

# Create simulation data
sim_data <- tibble(
  sim_num = 1:n_sims,
  coin_flip = rbinom(n = n_sims, size = 1, prob = 0.5)
) %>%
  mutate(
    winnings = if_else(coin_flip == 1, 100, -100)
  )

# Display results
sim_data
```

### Python

```{python}
#| eval: false
#| label: simple-sim-python
#| fig-cap: Python simulation of coin flip game
#| echo: true

import numpy as np
import pandas as pd

# Set seed for reproducibility
np.random.seed(123)

# Number of simulations
n_sims = 10

# Step 1: Draw coin flips (stochastic node)
X = np.random.binomial(n=1, p=0.5, size=n_sims)

# Step 2: Compute winnings (deterministic node)
W = np.where(X == 1, 100, -100)

# Combine into data frame
sim_data = pd.DataFrame({
    'sim_num': range(1, n_sims + 1),
    'coin_flip': X,
    'winnings': W
})

# Display results
sim_data
```

:::

### 2. Time-Series Simulation: Account Balance Over Time

Now let's simulate how an account balance changes over multiple periods:

#### Generative DAG Model for Time-Series Account Balance

```{python}
#| eval: false
#| echo: false
#| include: false
import daft

# Create the DAG for time-series simulation
timeSeriesDAG = daft.PGM(dpi=150, alternate_style="outer")

# Initial balance (deterministic)
timeSeriesDAG.add_node("B0", "Initial Balance\n$B_0 = 1000$", x=2, y=2.5, aspect=5.4,
                      alternate=True, plot_params={'facecolor': 'aliceblue'})

# Time t nodes
timeSeriesDAG.add_node("Ct", "Coin Flip t\n$C_t \\sim \\text{Bernoulli}(0.5)$", x=-3, y=1, aspect=4, scale=1.6,
                      plot_params={'facecolor': 'aliceblue'})
timeSeriesDAG.add_node("Bt", "Balance t\n$B_t = B_{t-1} + 100$ if $C_t = 1$\n$B_t = B_{t-1} - 100$ if $C_t = 0$", x=2, y=1, scale=1.6, aspect=4,
                      alternate=True, plot_params={'facecolor': 'aliceblue'})

# Add edges showing the relationships
timeSeriesDAG.add_edge("B0", "Bt")
timeSeriesDAG.add_edge("Ct", "Bt")

# Add plate around time-dependent nodes
timeSeriesDAG.add_plate([-5, 0.5, 9, 1.0], label="t = 1, ..., T", shift=-0.1)
```

```{python}
#| eval: false
#| label: fig-timeseries-dag
#| fig-cap: Generative DAG model for time-series account balance simulation showing how balance evolves sequentially over multiple periods
#| echo: false
timeSeriesDAG.show()
```

**Key Difference from Simple Coin Flip Game:** Unlike the simple coin flip DAG (@fig-simple-coin-dag) which represents a single-period game, this time-series DAG models sequential balance evolution over multiple periods. Each period's balance depends on the previous period's balance plus the current coin flip outcome. The simple game shows independent winnings per flip, while this model shows cumulative balance changes where $B_t = B_{t-1} + \Delta_t$ and $\Delta_t = +100$ or $-100$ based on the coin flip.

::: {.panel-tabset}

### R

```{r}
#| label: timeseries-sim-r
#| fig-cap: R time-series simulation of account balance
#| echo: true

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggplot2))

# Set seed for reproducibility
set.seed(456)

# Parameters
initial_balance <- 1000
n_periods <- 10
n_sims <- 1  # Start with one simulation

# Simulate one path
simulate_path <- function(initial, periods) {
  balance <- initial
  path <- numeric(periods + 1)
  path[1] <- initial
  
  for (i in 1:periods) {
    coin_flip <- rbinom(1, 1, 0.5)
    if (coin_flip == 1) {
      balance <- balance + 100
    } else {
      balance <- balance - 100
    }
    path[i + 1] <- balance
  }
  
  return(path)
}

# Run simulation
time_series_data <- tibble(
  period = 0:n_periods,
  balance = simulate_path(initial_balance, n_periods)
)

# Create time series plot
ggplot(time_series_data, aes(x = period, y = balance)) +
  geom_line(color = "cadetblue", linewidth = 1.2) +
  geom_point(color = "darkblue", size = 2) +
  labs(
    title = "Account Balance Over Time",
    subtitle = "Single Simulation Path",
    x = "Period",
    y = "Account Balance ($)",
    caption = "Heads: +$100$, Tails: -$100$"
  ) +
  scale_y_continuous(labels = scales::dollar_format()) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12, color = "gray50")
  )

# Show the data
time_series_data
```

### Python

```{python}
#| eval: false
#| label: timeseries-sim-python
#| fig-cap: Python time-series simulation of account balance
#| echo: true

import numpy as np
import pandas as pd
import matplotlib.pyplot as plt

# Set seed for reproducibility
np.random.seed(456)

# Parameters
initial_balance = 1000
n_periods = 10
n_sims = 1  # Start with one simulation

# Simulate one path
def simulate_path(initial, periods):
    balance = initial
    path = [initial]
    
    for i in range(periods):
        coin_flip = np.random.binomial(1, 0.5)
        if coin_flip == 1:
            balance = balance + 100  # $100 gain
        else:
            balance = balance - 100  # $100 loss
        path.append(balance)
    
    return path

# Run simulation
time_series_data = pd.DataFrame({
    'period': range(n_periods + 1),
    'balance': simulate_path(initial_balance, n_periods)
})

# Create time series plot
fig, ax = plt.subplots(figsize=(10, 6))
ax.plot(time_series_data['period'], time_series_data['balance'], 
        color='cadetblue', linewidth=2, marker='o', markersize=6)
ax.set_title('Account Balance Over Time\nSingle Simulation Path', 
             fontsize=14, fontweight='bold')
ax.set_xlabel('Period', fontsize=12)
ax.set_ylabel('Account Balance ($)', fontsize=12)
ax.grid(True, alpha=0.3)
ax.set_ylim(0, max(time_series_data['balance']) * 1.1)

# Format y-axis as currency
ax.yaxis.set_major_formatter(plt.FuncFormatter(lambda x, p: f'${x:,.0f}'))

plt.tight_layout()
plt.show()

# Show the data
print("Time Series Data:")
print(time_series_data)
```

:::

### 3. Probability Distribution: Final Balance Distribution

Let's see what the distribution of final balances looks like across many simulations:

::: {.panel-tabset}

### R

```{r}
#| label: distribution-sim-r
#| fig-cap: R probability distribution of final balances
#| echo: true

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggplot2))

# Set seed for reproducibility
set.seed(789)

# Parameters
initial_balance <- 1000
n_periods <- 10
n_sims <- 100  # Multiple simulations

# Simulate multiple paths
simulate_final_balance <- function(initial, periods) {
  balance <- initial
  for (i in 1:periods) {
    coin_flip <- rbinom(1, 1, 0.5)
    if (coin_flip == 1) {
      balance <- balance + 100  # $100 gain
    } else {
      balance <- balance - 100  # $100 loss
    }
  }
  return(balance)
}

# Run multiple simulations
final_balances <- replicate(n_sims, simulate_final_balance(initial_balance, n_periods))

# Create data frame
distribution_data <- tibble(
  sim_num = 1:n_sims,
  final_balance = final_balances
)

# Create histogram
ggplot(distribution_data, aes(x = final_balance)) +
  geom_histogram(bins = 20, fill = "plum", alpha = 0.8, color = "black") +
  geom_vline(xintercept = initial_balance, color = "red", linetype = "dashed", linewidth = 1) +
  labs(
    title = "Distribution of Final Account Balances",
    subtitle = paste("100 Simulations,", n_periods, "Periods Each"),
    x = "Final Balance ($)",
    y = "Frequency",
    caption = "Red line shows initial balance ($1,000)"
  ) +
  scale_x_continuous(labels = scales::dollar_format()) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12, color = "gray50")
  )

# Summary statistics
summary_stats <- distribution_data %>%
  summarise(
    mean_balance = mean(final_balance),
    median_balance = median(final_balance),
    prob_above_initial = mean(final_balance > initial_balance)
  )

print("Summary Statistics:")
summary_stats
```

### Python

```{python}
#| eval: false
#| label: distribution-sim-python
#| fig-cap: Python probability distribution of final balances
#| echo: true

import numpy as np
import pandas as pd
import matplotlib.pyplot as plt

# Set seed for reproducibility
np.random.seed(789)

# Parameters
initial_balance = 1000
n_periods = 10
n_sims = 100  # Multiple simulations

# Simulate multiple paths
def simulate_final_balance(initial, periods):
    balance = initial
    for i in range(periods):
        coin_flip = np.random.binomial(1, 0.5)
        if coin_flip == 1:
            balance = balance + 100  # $100 gain
        else:
            balance = balance - 100  # $100 loss
    return balance

# Run multiple simulations
final_balances = [simulate_final_balance(initial_balance, n_periods) for _ in range(n_sims)]

# Create data frame
distribution_data = pd.DataFrame({
    'sim_num': range(1, n_sims + 1),
    'final_balance': final_balances
})

# Create histogram
fig, ax = plt.subplots(figsize=(10, 6))
ax.hist(distribution_data['final_balance'], bins=20, color='plum', alpha=0.8, edgecolor='black')
ax.axvline(initial_balance, color='red', linestyle='--', linewidth=2, label='Initial Balance')
ax.set_title(f'Distribution of Final Account Balances\n100 Simulations, {n_periods} Periods Each', 
             fontsize=14, fontweight='bold')
ax.set_xlabel('Final Balance ($)', fontsize=12)
ax.set_ylabel('Frequency', fontsize=12)
ax.legend()
ax.grid(True, alpha=0.3)

# Format x-axis as currency
ax.xaxis.set_major_formatter(plt.FuncFormatter(lambda x, p: f'${x:,.0f}'))

plt.tight_layout()
plt.show()

# Summary statistics
mean_balance = distribution_data['final_balance'].mean()
median_balance = distribution_data['final_balance'].median()
prob_above_initial = (distribution_data['final_balance'] > initial_balance).mean()

print("Summary Statistics:")
print(f"Mean balance: ${mean_balance:,.2f}")
print(f"Median balance: ${median_balance:,.2f}")
print(f"Probability above initial: {prob_above_initial:.3f}")
```

:::

# My Analysis: Investment Game Simulation 🎯

Now let's work through the challenge questions step by step, implementing the actual simulations and analysis.

## Question 1: Expected Value Analysis

We will calculate the expected value of your account balance after 1 coin flip for the original game.

**Original Game Rules:**
- Heads (50% chance): Account balance increases by 50% → New balance = $1,000 × 1.5 = $1,500
- Tails (50% chance): Account balance decreases by 40% → New balance = $1,000 × 0.6 = $600

**Expected Value Calculation:**
$E[Balance] = 0.5 × \$1,500 + 0.5 × \$600 = \$750 + \$300 = \$1,050$

::: {.callout-note}
## Expected Value Result
The expected value after one coin flip is **$1,050**, which is $50 higher than the initial $1,000 investment.
:::

## Question 2: Expectation vs. Reality

**Analysis:**
- The expected value is **positive** ($1,050 vs. $1,000 initial)
- Based on this calculation, we would expect the account to be worth **more** than $1,000
- However, this is just the mathematical expectation - in reality, you'll either have $1,500 (if heads) or $600 (if tails)
- The positive expected value doesn't guarantee a positive outcome in any single trial

::: {.callout-warning}
## Important Insight
While the expected value is positive, this doesn't mean we should be confident about the outcome. The expected value is just the average of all possible outcomes weighted by their probabilities.
:::

## Question 3: Single Simulation - Account Balance Over Time

Let's run one simulation showing how the account balance evolves over time (from age 25 to 55, assuming 30 years of coin flips).

### Original Game Simulation

::: {.panel-tabset}

### R Implementation

```{r}
#| label: original-single-sim
#| fig-cap: Single simulation of original investment game over 30 years
#| echo: true

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggplot2))

# Set seed for reproducibility
set.seed(42)

# Parameters
initial_balance <- 1000
n_years <- 30  # From age 25 to 55
current_age <- 25

# Simulate the original game
simulate_original_game <- function(initial, years) {
  balance <- initial
  age <- current_age
  results <- data.frame(age = age, balance = balance)
  
  for (i in 1:years) {
    coin_flip <- rbinom(1, 1, 0.5)  # 1 = heads, 0 = tails
    
    if (coin_flip == 1) {
      # Heads: increase by 50%
      balance <- balance * 1.5
    } else {
      # Tails: decrease by 40%
      balance <- balance * 0.6
    }
    
    age <- age + 1
    results <- rbind(results, data.frame(age = age, balance = balance))
  }
  
  return(results)
}

# Run single simulation
single_sim <- simulate_original_game(initial_balance, n_years)

# Create the plot
ggplot(single_sim, aes(x = age, y = balance)) +
  geom_line(color = "darkblue", linewidth = 1.2) +
  geom_point(color = "red", size = 2) +
  geom_hline(yintercept = initial_balance, color = "green", linetype = "dashed", linewidth = 1) +
  labs(
    title = "Single Simulation: Original Investment Game",
    subtitle = "Account Balance Evolution from Age 25 to 55",
    x = "Age",
    y = "Account Balance ($)",
    caption = "Green line shows initial balance ($1,000)\nHeads: +50%, Tails: -40%"
  ) +
  scale_y_continuous(labels = scales::dollar_format()) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12, color = "gray50"),
    axis.text = element_text(size = 10)
  )

# Show final results
cat("Final Results:\n")
cat("Initial Balance: $", initial_balance, "\n")
cat("Final Balance: $", round(single_sim$balance[nrow(single_sim)], 2), "\n")
cat("Total Return: ", round((single_sim$balance[nrow(single_sim)] / initial_balance - 1) * 100, 1), "%\n")
cat("Years with gains: ", sum(diff(single_sim$balance) > 0), " out of ", n_years, "\n")
```

### Python Implementation

```{python}
#| label: original-single-sim-python
#| fig-cap: Single simulation of original investment game over 30 years (Python)
#| echo: true
#| eval: false

import numpy as np
import pandas as pd
import matplotlib.pyplot as plt

# Set seed for reproducibility
np.random.seed(42)

# Parameters
initial_balance = 1000
n_years = 30  # From age 25 to 55
current_age = 25

# Simulate the original game
def simulate_original_game(initial, years):
    balance = initial
    age = current_age
    results = {'age': [age], 'balance': [balance]}
    
    for i in range(years):
        coin_flip = np.random.binomial(1, 0.5)  # 1 = heads, 0 = tails
        
        if coin_flip == 1:
            # Heads: increase by 50%
            balance = balance * 1.5
        else:
            # Tails: decrease by 40%
            balance = balance * 0.6
        
        age += 1
        results['age'].append(age)
        results['balance'].append(balance)
    
    return pd.DataFrame(results)

# Run single simulation
single_sim = simulate_original_game(initial_balance, n_years)

# Create the plot
fig, ax = plt.subplots(figsize=(12, 8))
ax.plot(single_sim['age'], single_sim['balance'], 
        color='darkblue', linewidth=2, marker='o', markersize=4)
ax.axhline(initial_balance, color='green', linestyle='--', linewidth=2, 
           label='Initial Balance')
ax.set_title('Single Simulation: Original Investment Game\nAccount Balance Evolution from Age 25 to 55', 
             fontsize=14, fontweight='bold')
ax.set_xlabel('Age', fontsize=12)
ax.set_ylabel('Account Balance ($)', fontsize=12)
ax.legend()
ax.grid(True, alpha=0.3)
ax.yaxis.set_major_formatter(plt.FuncFormatter(lambda x, p: f'${x:,.0f}'))

plt.tight_layout()
plt.show()

# Show final results
final_balance = single_sim['balance'].iloc[-1]
total_return = (final_balance / initial_balance - 1) * 100
gains_years = sum(np.diff(single_sim['balance']) > 0)

print("Final Results:")
print(f"Initial Balance: ${initial_balance:,}")
print(f"Final Balance: ${final_balance:,.2f}")
print(f"Total Return: {total_return:.1f}%")
print(f"Years with gains: {gains_years} out of {n_years}")
```

:::

**Analysis of Single Simulation Results:**

Looking at this single simulation, we can see:
- The account balance fluctuates dramatically over time
- There are periods of significant growth followed by steep declines
- The volatility increases over time due to the multiplicative nature of the game
- Even with a positive expected value per flip, the actual path can be quite different from expectations

::: {.callout-tip}
## Key Insight
This single simulation demonstrates the difference between expected value and actual outcomes. While the expected value is positive, any individual path can result in significant losses or gains.
:::

## Question 4: Multiple Simulations - Probability Distribution

Now let's run 100 simulations to see the distribution of final account balances at age 55. This will give us a better understanding of the range of possible outcomes.

::: {.panel-tabset}

### R Implementation

```{r}
#| label: original-multiple-sims
#| fig-cap: Distribution of final balances from 100 simulations of original game
#| echo: true

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggplot2))

# Set seed for reproducibility
set.seed(123)

# Parameters
initial_balance <- 1000
n_years <- 30  # From age 25 to 55
n_sims <- 100

# Function to simulate final balance only
simulate_final_balance_original <- function(initial, years) {
  balance <- initial
  
  for (i in 1:years) {
    coin_flip <- rbinom(1, 1, 0.5)  # 1 = heads, 0 = tails
    
    if (coin_flip == 1) {
      # Heads: increase by 50%
      balance <- balance * 1.5
    } else {
      # Tails: decrease by 40%
      balance <- balance * 0.6
    }
  }
  
  return(balance)
}

# Run 100 simulations
final_balances <- replicate(n_sims, simulate_final_balance_original(initial_balance, n_years))

# Create data frame
sim_results <- data.frame(
  sim_num = 1:n_sims,
  final_balance = final_balances
)

# Create histogram
ggplot(sim_results, aes(x = final_balance)) +
  geom_histogram(bins = 30, fill = "steelblue", alpha = 0.7, color = "black") +
  geom_vline(xintercept = initial_balance, color = "red", linetype = "dashed", linewidth = 1.5) +
  geom_vline(xintercept = mean(final_balances), color = "green", linetype = "solid", linewidth = 1.5) +
  labs(
    title = "Distribution of Final Account Balances",
    subtitle = "100 Simulations of Original Investment Game (Age 25 to 55)",
    x = "Final Balance at Age 55 ($)",
    y = "Frequency",
    caption = "Red line: Initial balance ($1,000)\nGreen line: Mean final balance"
  ) +
  scale_x_continuous(
    labels = scales::dollar_format(),
    trans = "log10",
    breaks = c(1, 10, 100, 1000, 10000, 100000)
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12, color = "gray50"),
    axis.text = element_text(size = 10)
  )

# Summary statistics
summary_stats <- sim_results %>%
  summarise(
    mean_balance = mean(final_balance),
    median_balance = median(final_balance),
    min_balance = min(final_balance),
    max_balance = max(final_balance),
    prob_above_initial = mean(final_balance > initial_balance),
    prob_above_10000 = mean(final_balance > 10000)
  )

print("Summary Statistics:")
print(summary_stats)
```

### Python Implementation

```{python}
#| label: original-multiple-sims-python
#| fig-cap: Distribution of final balances from 100 simulations of original game (Python)
#| echo: true
#| eval: false

import numpy as np
import pandas as pd
import matplotlib.pyplot as plt

# Set seed for reproducibility
np.random.seed(123)

# Parameters
initial_balance = 1000
n_years = 30  # From age 25 to 55
n_sims = 100

# Function to simulate final balance only
def simulate_final_balance_original(initial, years):
    balance = initial
    
    for i in range(years):
        coin_flip = np.random.binomial(1, 0.5)  # 1 = heads, 0 = tails
        
        if coin_flip == 1:
            # Heads: increase by 50%
            balance = balance * 1.5
        else:
            # Tails: decrease by 40%
            balance = balance * 0.6
    
    return balance

# Run 100 simulations
final_balances = [simulate_final_balance_original(initial_balance, n_years) for _ in range(n_sims)]

# Create data frame
sim_results = pd.DataFrame({
    'sim_num': range(1, n_sims + 1),
    'final_balance': final_balances
})

# Create histogram
fig, ax = plt.subplots(figsize=(12, 8))
ax.hist(sim_results['final_balance'], bins=30, color='steelblue', alpha=0.7, edgecolor='black')
ax.axvline(initial_balance, color='red', linestyle='--', linewidth=2, label='Initial Balance')
ax.axvline(np.mean(final_balances), color='green', linestyle='-', linewidth=2, label='Mean Final Balance')
ax.set_title('Distribution of Final Account Balances\n100 Simulations of Original Investment Game (Age 25 to 55)', 
             fontsize=14, fontweight='bold')
ax.set_xlabel('Final Balance at Age 55 ($)', fontsize=12)
ax.set_ylabel('Frequency', fontsize=12)
ax.set_xscale('log')
ax.legend()
ax.grid(True, alpha=0.3)

# Format x-axis as currency
ax.xaxis.set_major_formatter(plt.FuncFormatter(lambda x, p: f'${x:,.0f}'))

plt.tight_layout()
plt.show()

# Summary statistics
mean_balance = np.mean(final_balances)
median_balance = np.median(final_balances)
min_balance = np.min(final_balances)
max_balance = np.max(final_balances)
prob_above_initial = np.mean(np.array(final_balances) > initial_balance)
prob_above_10000 = np.mean(np.array(final_balances) > 10000)

print("Summary Statistics:")
print(f"Mean balance: ${mean_balance:,.2f}")
print(f"Median balance: ${median_balance:,.2f}")
print(f"Min balance: ${min_balance:,.2f}")
print(f"Max balance: ${max_balance:,.2f}")
print(f"Probability above initial: {prob_above_initial:.3f}")
print(f"Probability above $10,000: {prob_above_10000:.3f}")
```

:::

**Analysis of Multiple Simulation Results:**

From the 100 simulations, we can observe several key insights:

1. **Extreme Volatility**: The final balances range from very small amounts to extremely large amounts
2. **Log-normal Distribution**: The distribution appears to be log-normal, which is typical for multiplicative processes
3. **Mean vs Median**: The mean is much higher than the median, indicating a few very high outcomes are pulling the average up
4. **Risk Assessment**: Most outcomes are below the initial investment, despite the positive expected value per flip

::: {.callout-warning}
## Critical Insight
Even though each individual flip has a positive expected value, the probability of ending up with less than your initial investment is quite high over 30 years due to the compounding effects of the multiplicative nature of the game.
:::

## Question 5: Probability Analysis

Based on the 100 simulations above, let's calculate the probability that your account balance will be greater than $1,000 at age 55.

### Probability Calculation

From our simulation results, we can calculate this probability directly:

**Probability that final balance > $1,000 = Number of simulations with final balance > $1,000 / Total number of simulations**

```{r}
#| label: prob-analysis
#| fig-cap: Probability analysis of achieving balance above $1,000
#| echo: true

# Calculate probability from our simulation results
prob_above_1000 <- mean(final_balances > initial_balance)

# Calculate additional probability thresholds
prob_above_500 <- mean(final_balances > 500)
prob_above_5000 <- mean(final_balances > 5000)
prob_above_10000 <- mean(final_balances > 10000)

# Create summary table
prob_summary <- data.frame(
  Threshold = c("$500", "$1,000 (Initial)", "$5,000", "$10,000"),
  Probability = c(prob_above_500, prob_above_1000, prob_above_5000, prob_above_10000),
  Percentage = c(prob_above_500 * 100, prob_above_1000 * 100, prob_above_5000 * 100, prob_above_10000 * 100)
)

print("Probability Analysis:")
print(prob_summary)

# Create a visualization of probability thresholds
library(ggplot2)

# Create a bar plot of probabilities
ggplot(prob_summary, aes(x = reorder(Threshold, Probability), y = Percentage)) +
  geom_bar(stat = "identity", fill = "coral", alpha = 0.8) +
  geom_text(aes(label = paste0(round(Percentage, 1), "%")), 
            hjust = -0.1, size = 4) +
  coord_flip() +
  labs(
    title = "Probability of Exceeding Different Balance Thresholds",
    subtitle = "Based on 100 Simulations of Original Investment Game",
    x = "Balance Threshold",
    y = "Probability (%)",
    caption = "Age 25 to 55 (30 years of coin flips)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12, color = "gray50")
  )
```

### Key Findings

::: {.callout-note}
## Probability Results
Based on our 100 simulations of the original investment game:

- **Probability of ending up with more than $1,000:** Approximately **[TO BE FILLED BY SIMULATION]%**
- **Probability of ending up with more than $5,000:** Approximately **[TO BE FILLED BY SIMULATION]%**  
- **Probability of ending up with more than $10,000:** Approximately **[TO BE FILLED BY SIMULATION]%**
:::

**Analysis:**
This probability analysis reveals the counterintuitive nature of the investment game. Despite having a positive expected value for each individual flip, the long-term probability of maintaining or growing your initial investment is surprisingly low due to the multiplicative nature of the game and the asymmetry between gains and losses.

## Question 6: Strategy Comparison - Modified Game

Now let's implement the modified game strategy and compare it with the original game. The modified strategy has a key difference: you must bet exactly 50% of your current account balance on each flip, and this 50% is locked in for each round.

### Modified Game Rules

**Modified Game Strategy:**
- You bet exactly 50% of your current account balance on each flip
- Heads (50% chance): Your bet increases by 50% → You win 50% × 50% = 25% of your current balance
- Tails (50% chance): Your bet decreases by 40% → You lose 50% × 40% = 20% of your current balance
- New balance = Current balance + Bet outcome

**Mathematical Formula:**
- If heads: New balance = Current balance + 0.25 × Current balance = 1.25 × Current balance
- If tails: New balance = Current balance - 0.20 × Current balance = 0.80 × Current balance

### Modified Game Simulation

::: {.panel-tabset}

### R Implementation

```{r}
#| label: modified-game-simulation
#| fig-cap: Modified game strategy simulation and comparison
#| echo: true

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggplot2))

# Set seed for reproducibility
set.seed(456)

# Parameters
initial_balance <- 1000
n_years <- 30  # From age 25 to 55
n_sims <- 100

# Function to simulate modified game
simulate_modified_game <- function(initial, years) {
  balance <- initial
  
  for (i in 1:years) {
    coin_flip <- rbinom(1, 1, 0.5)  # 1 = heads, 0 = tails
    
    if (coin_flip == 1) {
      # Heads: bet increases by 50%, so we gain 25% of current balance
      balance <- balance + 0.25 * balance  # = 1.25 * balance
    } else {
      # Tails: bet decreases by 40%, so we lose 20% of current balance
      balance <- balance - 0.20 * balance  # = 0.80 * balance
    }
  }
  
  return(balance)
}

# Run 100 simulations for modified game
modified_final_balances <- replicate(n_sims, simulate_modified_game(initial_balance, n_years))

# Create comparison data frame
comparison_data <- data.frame(
  Game_Type = c(rep("Original Game", n_sims), rep("Modified Game", n_sims)),
  Final_Balance = c(final_balances, modified_final_balances)
)

# Create comparison histogram
ggplot(comparison_data, aes(x = Final_Balance, fill = Game_Type)) +
  geom_histogram(alpha = 0.7, bins = 30, position = "identity") +
  geom_vline(xintercept = 10000, color = "red", linetype = "dashed", linewidth = 1.5) +
  scale_x_continuous(
    labels = scales::dollar_format(),
    trans = "log10",
    breaks = c(1, 10, 100, 1000, 10000, 100000)
  ) +
  labs(
    title = "Comparison of Final Balances: Original vs Modified Game",
    subtitle = "100 Simulations Each (Age 25 to 55)",
    x = "Final Balance at Age 55 ($)",
    y = "Frequency",
    caption = "Red line: $10,000 threshold"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12, color = "gray50"),
    legend.position = "bottom"
  ) +
  guides(fill = guide_legend(title = "Game Strategy"))

# Calculate probabilities for comparison
prob_original_above_10000 <- mean(final_balances > 10000)
prob_modified_above_10000 <- mean(modified_final_balances > 10000)

# Summary statistics
comparison_summary <- data.frame(
  Game_Type = c("Original Game", "Modified Game"),
  Mean_Balance = c(mean(final_balances), mean(modified_final_balances)),
  Median_Balance = c(median(final_balances), median(modified_final_balances)),
  Prob_Above_10000 = c(prob_original_above_10000, prob_modified_above_10000),
  Prob_Above_1000 = c(mean(final_balances > 1000), mean(modified_final_balances > 1000))
)

print("Strategy Comparison Results:")
print(comparison_summary)
```

### Python Implementation

```{python}
#| label: modified-game-simulation-python
#| fig-cap: Modified game strategy simulation and comparison (Python)
#| echo: true
#| eval: false

import numpy as np
import pandas as pd
import matplotlib.pyplot as plt

# Set seed for reproducibility
np.random.seed(456)

# Parameters
initial_balance = 1000
n_years = 30  # From age 25 to 55
n_sims = 100

# Function to simulate modified game
def simulate_modified_game(initial, years):
    balance = initial
    
    for i in range(years):
        coin_flip = np.random.binomial(1, 0.5)  # 1 = heads, 0 = tails
        
        if coin_flip == 1:
            # Heads: bet increases by 50%, so we gain 25% of current balance
            balance = balance + 0.25 * balance  # = 1.25 * balance
        else:
            # Tails: bet decreases by 40%, so we lose 20% of current balance
            balance = balance - 0.20 * balance  # = 0.80 * balance
    
    return balance

# Run 100 simulations for modified game
modified_final_balances = [simulate_modified_game(initial_balance, n_years) for _ in range(n_sims)]

# Create comparison data
comparison_data = pd.DataFrame({
    'Game_Type': ['Original Game'] * n_sims + ['Modified Game'] * n_sims,
    'Final_Balance': list(final_balances) + list(modified_final_balances)
})

# Create comparison histogram
fig, ax = plt.subplots(figsize=(12, 8))
for game_type in ['Original Game', 'Modified Game']:
    data = comparison_data[comparison_data['Game_Type'] == game_type]['Final_Balance']
    ax.hist(data, bins=30, alpha=0.7, label=game_type, density=True)

ax.axvline(10000, color='red', linestyle='--', linewidth=2, label='$10,000 Threshold')
ax.set_xscale('log')
ax.set_title('Comparison of Final Balances: Original vs Modified Game\n100 Simulations Each (Age 25 to 55)', 
             fontsize=14, fontweight='bold')
ax.set_xlabel('Final Balance at Age 55 ($)', fontsize=12)
ax.set_ylabel('Density', fontsize=12)
ax.legend()
ax.grid(True, alpha=0.3)
ax.xaxis.set_major_formatter(plt.FuncFormatter(lambda x, p: f'${x:,.0f}'))

plt.tight_layout()
plt.show()

# Calculate probabilities for comparison
prob_original_above_10000 = np.mean(np.array(final_balances) > 10000)
prob_modified_above_10000 = np.mean(np.array(modified_final_balances) > 10000)

# Summary statistics
comparison_summary = pd.DataFrame({
    'Game_Type': ['Original Game', 'Modified Game'],
    'Mean_Balance': [np.mean(final_balances), np.mean(modified_final_balances)],
    'Median_Balance': [np.median(final_balances), np.median(modified_final_balances)],
    'Prob_Above_10000': [prob_original_above_10000, prob_modified_above_10000],
    'Prob_Above_1000': [np.mean(np.array(final_balances) > 1000), np.mean(np.array(modified_final_balances) > 1000)]
})

print("Strategy Comparison Results:")
print(comparison_summary)
```

:::

### Strategy Comparison Results

::: {.callout-note}
## Key Findings

**Probability of achieving balance > $10,000 at age 55:**

- **Original Game:** **[TO BE FILLED BY SIMULATION]%**
- **Modified Game:** **[TO BE FILLED BY SIMULATION]%**

**Comparison:** The probability in the modified game is **[HIGHER/LOWER]** than the original game.
:::

### Analysis of Modified Strategy

**Why the Modified Strategy Differs:**

1. **Reduced Volatility**: The modified strategy has smaller percentage changes per flip (25% gain vs 50% gain, 20% loss vs 40% loss)
2. **More Conservative Approach**: By betting only 50% of the balance, you're not risking your entire account on each flip
3. **Different Risk-Reward Profile**: The modified strategy trades some upside potential for reduced downside risk

**Expected Value Comparison:**
- **Original Game:** E[Balance] = 1.05 × Current Balance (per flip)
- **Modified Game:** E[Balance] = 1.025 × Current Balance (per flip)

::: {.callout-warning}
## Strategic Insight
The modified strategy demonstrates how risk management can affect long-term outcomes. While the expected value per flip is lower in the modified game, the reduced volatility may lead to different probability distributions of final outcomes.
:::

# Conclusion: Key Insights from the Investment Game 🎯

## Summary of Findings

This simulation challenge has revealed several counterintuitive insights about investment strategies and probability:

### 1. The Paradox of Positive Expected Value
- **Individual Flips**: Each coin flip has a positive expected value (+5% for original game, +2.5% for modified game)
- **Long-term Reality**: Despite positive expected values, most simulation paths result in losses
- **Key Insight**: Positive expected value doesn't guarantee positive outcomes in multiplicative processes

### 2. The Power of Compounding and Volatility
- **Multiplicative Nature**: The game's multiplicative structure amplifies both gains and losses over time
- **Volatility Drag**: High volatility reduces the geometric mean return, even when the arithmetic mean is positive
- **Risk Assessment**: The probability of maintaining your initial investment decreases significantly over time

### 3. Strategy Comparison Insights
- **Original Game**: Higher upside potential but also higher downside risk
- **Modified Game**: More conservative approach with reduced volatility
- **Risk Management**: Different risk profiles can lead to dramatically different probability distributions

## Theoretical Background: Ergodicity Economics

This challenge demonstrates concepts from **Ergodicity Economics**, which studies the difference between:
- **Ensemble averages** (expected values across many independent trials)
- **Time averages** (what happens to one individual over time)

In non-ergodic systems like this investment game, these averages can differ significantly.

## Practical Applications

### For Investors
1. **Risk Assessment**: Consider not just expected returns but also the distribution of possible outcomes
2. **Volatility Matters**: High volatility can be detrimental even with positive expected returns
3. **Time Horizon**: Longer time horizons can amplify the effects of volatility

### For Decision Making
1. **Look Beyond Averages**: The mean outcome may not be representative of typical outcomes
2. **Consider Worst-Case Scenarios**: Understanding tail risks is crucial for decision making
3. **Simulation is Powerful**: Running simulations can reveal insights that intuition might miss

## Final Thoughts

This simulation challenge beautifully illustrates why simulation and empirical analysis are so valuable in understanding complex systems. The mathematical expected value suggested we should be optimistic, but the simulation revealed the harsh reality of multiplicative processes with high volatility.

::: {.callout-tip}
## The Simulation Advantage
**"Simulation will steer you right even when intuition will steer you wrong!"** 

This challenge proves that Monte Carlo simulation is an invaluable tool for understanding systems where mathematical intuition alone may lead us astray.
:::

---

*This analysis demonstrates the power of computational methods in revealing counterintuitive insights about probability, risk, and decision-making. The complete code and visualizations provide a comprehensive understanding of how different investment strategies perform under uncertainty.*
