---
title: "Simulation Challenge"
subtitle: "Generative Models and Monte Carlo Simulation"
format:
  html: default

execute:
  echo: false
  eval: true
knitr:
  opts_chunk:
    python.reticulate: false
---

# 🎲 Simulation Challenge - Monte Carlo Analysis

## Challenge Overview

**Your Mission:** Create a comprehensive Quarto document that simulates one or two investment strategies, analyzes the results, and demonstrates your ability to present counter-intuitive findings compellingly. Then render the document to HTML and deploy it via GitHub Pages from a new repository called "simulationChallenge."

::: {.callout-warning}
## ⚠️ AI Partnership Required

This challenge pushes boundaries intentionally. You'll tackle problems that normally require weeks of study, but with Cursor AI as your partner (and your brain keeping it honest), you can accomplish more than you thought possible.

**The new reality:** The four stages of competence are Ignorance → Awareness → Learning → Mastery. AI lets us produce Mastery-level work while operating primarily in the Awareness stage. I focus on awareness training, you leverage AI for execution, and together we create outputs that used to require years of dedicated study.
:::

## The Investment Game 🎯

### Original Game Strategy

::: {#exm-ErgodicityEconomicsExample}
Imagine you are offered the following game and given a $1,000 budget in a special account to play the game: I will flip a coin, and if it comes up heads, we increase your account's balance by 50%; if it comes up tails, we reduce your account's balance by 40%. We are not only doing this once, but we will do it once per year until you turn 55. When you turn 55, you will receive the balance in your account.
:::



## Challenge Requirements 📋

### Minimum Requirements for Any Points on Challenge

1. **Create a Quarto Document:** Write a concise quarto markdown file that includes a narrative of what you are doing along with the requested code, results, and visualizations of your simulations.

   ::: {.callout-tip}
   ## 💡 Pro Tip: Source File Reference
   
   **Need help with Quarto syntax or DAFT diagrams?** You can find the complete source `.qmd` file for this challenge at:
   
   **GitHub Repository:** [flyaflya/buad442Fall2025/challenges/03-Simulation-Challenge/simulationChallenge.qmd](https://github.com/flyaflya/buad442Fall2025/blob/main/challenges/03-Simulation-Challenge/simulationChallenge.qmd)
   
   This is a great resource for understanding Quarto syntax, seeing how DAFT diagrams are implemented, and learning from the complete working example!
   :::

2. **Render to HTML:** You must render the quarto markdown file to HTML.

3. **GitHub Repository:** The rendered HTML must be uploaded to a new GitHub repository called "simulationChallenge" in your Github account.

4. **GitHub Pages Setup:** The repository should be made the source of your github pages:

   - Go to your repository settings (click the "Settings" tab in your GitHub repository)
   - Scroll down to the "Pages" section in the left sidebar
   - Under "Source", select "Deploy from a branch"
   - Choose "main" branch and "/ (root)" folder
   - Click "Save"
   - Your site will be available at: `https://[your-username].github.io/simulationChallenge/`
   - **Note:** It may take a few minutes for the site to become available after enabling Pages

## Grading Rubric 🎓

### Questions to Answer for 75% Grade on Challenge

1. **Expected Value Analysis:** What is the "expected value" of your account balance after 1 coin flip for the original game?

2. **Expectation vs. Reality:** Is the expected value positive or negative? Do you expect your account to be worth more or less than $1,000 based on this result?

3. **Single Simulation:** Run one simulation showing the dynamics of your account balance over time. Make an object-oriented matplotlib OR ggplot2 plot showing your simulated account balance over time (i.e. as you age). Comment on the results, are you happy?

### Questions to Answer for 85% Grade on Challenge

4. **Multiple Simulations:** Run 100 simulations modelling the dynamics of your account balance over time. Make an object-oriented matplotlib OR ggplot2 plot showing a probability distribution of the 100 simulatedaccount balance at age 55. Comment on the results, are you happy? Why or why not?

### Questions to Answer for 95% Grade on Challenge

5. **Probability Analysis:** Based on the 100 simulations above, what is the probability that your account balance will be greater than $1,000 at age 55?

### Questions to Answer for 100% Grade on Challenge

6. **Strategy Comparison:** Run 100 simulations for the modified game strategy shown below in @exm-ErgodicityEconomicsExampleModified. What is the probability that your account balance will be greater than $10,000 at age 55? Is this probability higher or lower than the probability in the original game?

### Modified Game Strategy

::: {#exm-ErgodicityEconomicsExampleModified}
Imagine you are offered the following game and given a $1,000 budget in a special account to play the game: I will flip a coin, and if it comes up heads, we increase your bet by 50%; if it comes up tails, we reduce your bet by 40%. You must bet exactly 50% of your current account balance on each flip, and this 50% is locked in for each round. We are not only doing this once, but we will do it once per year until you turn 55. When you turn 55, you will receive the balance in your account.
:::

## Technical Implementation Preferences 💡

### Setting Up Your Analysis

**For R Users:**

- Use `tidyverse` for data manipulation
- Use `ggplot2` for visualizations
- Use `set.seed()` for reproducible results

**For Python Users:**

- Use `numpy` for numerical operations
- Use `pandas` for data manipulation
- Use `matplotlib` (object-oriented)
- Use `np.random.seed()` for reproducible results

### Visualization Preferences

- **Professional Styling:** Use consistent colors, clear labels, readable fonts, and informative titles

## Submission Checklist ✅

**Minimum Requirements (Required for Any Points):**

- [ ] Quarto document created with clear narrative
- [ ] Document rendered to HTML successfully
- [ ] Repository "simulationChallenge" created
- [ ] HTML files uploaded to repository
- [ ] GitHub Pages enabled and working
- [ ] Site accessible at `https://[your-username].github.io/simulationChallenge/`

**75% Grade Requirements:**

- [ ] Expected value calculations shown (Question 1)
- [ ] Expectation vs. reality analysis (Question 2)
- [ ] Single simulation with time series plot (Question 3)
- [ ] Clear interpretation of single simulation results

**85% Grade Requirements:**

- [ ] 100 simulations with distribution analysis (Question 4)
- [ ] Probability distribution plot of final account balances
- [ ] Clear interpretation of multiple simulation results

**95% Grade Requirements:**

- [ ] Probability calculations for original strategy (Question 5)
- [ ] Analysis of probability that balance > $1,000 at age 55

**100% Grade Requirements:**

- [ ] 100 simulations for modified strategy (Question 6)
- [ ] Probability calculations for modified strategy
- [ ] Comparative analysis between both strategies
- [ ] Analysis of probability that balance > $10,000 at age 55

**Code Quality (All Grades):**

- [ ] Reproducible results (seeds set)
- [ ] Clean, well-commented code
- [ ] Appropriate use of functions and loops
- [ ] Professional visualization styling

### Resources

- **Quarto Markdown:** [quarto.org/docs/authoring/markdown-basics.html](https://quarto.org/docs/authoring/markdown-basics.html)
- **Quarto Documentation:** [quarto.org/docs](https://quarto.org/docs)
- **R for Data Science:** [r4ds.had.co.nz](https://r4ds.had.co.nz)
- **Python Data Science Handbook:** [jakevdp.github.io/PythonDataScienceHandbook](https://jakevdp.github.io/PythonDataScienceHandbook)

### Getting Started Tips

::: {.callout-note}
## 🎯 Navy SEALs Motto

> "Slow is Smooth and Smooth is Fast"

*Take your time to understand the simulation mechanics, plan your approach carefully, and execute with precision. Rushing through this challenge will only lead to errors and confusion.*
:::

- **Browse [Essential Simulation Concepts](@sec-simulation-concepts):** This section will give you a good understanding of the concepts you need to know to complete the challenge.
- **Start Simple:** Begin with a single simulation to understand the mechanics
- **Document Everything:** Explain your reasoning and interpret your results
- **Forgetting to Set Seeds:** Always set random seeds for reproducible results
- **Total time to complete:** ~3-4 hours for the 100% grade ⏱️
- **Good luck, and remember simulation will steer you right even when intuition will steer you wrong!** 🎲

::: {.callout-warning}
## 💾 Important: Save Your Work Frequently!

**Before you start coding:** Make sure to commit your work often using the Source Control panel in Cursor (Ctrl+Shift+G or Cmd+Shift+G). This prevents the AI from overwriting your progress and ensures you don't lose your work.

**Commit after each major step:**

- After completing each simulation example
- After finishing each challenge question
- Before asking the AI for help with new code

**How to commit:**

1. Open Source Control panel (Ctrl+Shift+G)
2. Stage your changes (+ button)
3. Write a descriptive commit message
4. Click the checkmark to commit

*Remember: Frequent commits are your safety net!*
:::


# My Analysis: Investment Game Simulation 🎯

Now let's work through the challenge questions step by step, implementing the actual simulations and analysis.

## Question 1: Expected Value After 1 Coin Flip

**Answer:** The expected value is **$1,050**

**Calculation:**
- Heads (50% chance): $1,000 × 1.5 = $1,500
- Tails (50% chance): $1,000 × 0.6 = $600
- Expected Value = 0.5 × $1,500 + 0.5 × $600 = **$1,050**

## Question 2: Expectation vs. Reality

**Answer:** The expected value is **positive** ($1,050 vs. $1,000), so we expect the account to be worth **more** than $1,000.

**However:** This is just math! In reality, you'll either get $1,500 or $600 - the expected value doesn't guarantee anything.

## Question 3: Single Simulation Over Time

```{r}
#| label: single-simulation
#| fig-cap: One simulation showing account balance from age 25 to 55
#| echo: false

set.seed(42)
initial_balance <- 1000
n_years <- 30

# Simulate 30 years
balance <- initial_balance
age <- 25
results <- data.frame(age = age, balance = balance)

for (i in 1:n_years) {
  coin_flip <- rbinom(1, 1, 0.5)  # 1 = heads, 0 = tails
  
  if (coin_flip == 1) {
    balance <- balance * 1.5  # +50%
  } else {
    balance <- balance * 0.6  # -40%
  }
  
  age <- age + 1
  results <- rbind(results, data.frame(age = age, balance = balance))
}

# Create ggplot2 plot
library(ggplot2)
ggplot(results, aes(x = age, y = balance)) +
  geom_line(color = "darkblue", linewidth = 1.2) +
  geom_point(color = "red", size = 2) +
  geom_hline(yintercept = initial_balance, color = "green", linetype = "dashed", linewidth = 1) +
  labs(
    title = "Single Simulation: Account Balance Over Time",
    subtitle = "Account Balance Evolution from Age 25 to 55",
    x = "Age",
    y = "Account Balance ($)",
    caption = "Green line shows initial balance ($1,000)\nHeads: +50%, Tails: -40%"
  ) +
  scale_y_continuous(labels = scales::dollar_format()) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12, color = "gray50"),
    axis.text = element_text(size = 10)
  )

# Show final result
final_balance <- results$balance[nrow(results)]
total_return <- round((final_balance / initial_balance - 1) * 100, 1)

cat("Final Balance at Age 55: $", round(final_balance, 2), "\n")
cat("Total Return: ", total_return, "%\n")
```

**Answer:** In this single simulation, the final balance was **$`r round(final_balance, 2)`** with a **`r total_return`%** return. 

**Am I happy?** Based on this result, I am **[happy/surprised]** because the simulation shows **[a significant gain/loss]** over 30 years.

## Question 4: 100 Simulations - Distribution of Results

```{r}
#| label: multiple-simulations
#| fig-cap: Distribution of final balances from 100 simulations
#| echo: false

set.seed(123)
initial_balance <- 1000
n_years <- 30
n_sims <- 100

# Function to run one complete simulation
simulate_final_balance <- function(initial, years) {
  balance <- initial
  
  for (i in 1:years) {
    coin_flip <- rbinom(1, 1, 0.5)  # 1 = heads, 0 = tails
    
    if (coin_flip == 1) {
      balance <- balance * 1.5  # +50%
    } else {
      balance <- balance * 0.6  # -40%
    }
  }
  
  return(balance)
}

# Run 100 simulations
final_balances <- replicate(n_sims, simulate_final_balance(initial_balance, n_years))

# Create histogram using ggplot2
library(ggplot2)
sim_results <- data.frame(final_balance = final_balances)
ggplot(sim_results, aes(x = final_balance)) +
  geom_histogram(bins = 20, fill = "steelblue", alpha = 0.7, color = "black") +
  geom_vline(xintercept = initial_balance, color = "red", linetype = "dashed", linewidth = 1.5) +
  labs(
    title = "Distribution of Final Balances (100 Simulations)",
    subtitle = "100 Simulations of Original Investment Game (Age 25 to 55)",
    x = "Final Balance at Age 55 ($)",
    y = "Frequency",
    caption = "Red line: Initial balance ($1,000)"
  ) +
  scale_x_continuous(
    labels = scales::dollar_format(),
    trans = "log10",
    breaks = c(1, 10, 100, 1000, 10000, 100000)
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12, color = "gray50"),
    axis.text = element_text(size = 10)
  )

# Summary statistics
prob_above_1000 <- mean(final_balances > 1000)
mean_balance <- mean(final_balances)
median_balance <- median(final_balances)

cat("Results from 100 simulations:\n")
cat("Mean final balance: $", round(mean_balance, 2), "\n")
cat("Median final balance: $", round(median_balance, 2), "\n")
cat("Probability above $1,000: ", round(prob_above_1000 * 100, 1), "%\n")
```

**Answer:** From 100 simulations:
- **Mean final balance:** $`r round(mean_balance, 2)`
- **Median final balance:** $`r round(median_balance, 2)`  
- **Probability above $1,000:** `r round(prob_above_1000 * 100, 1)`%

**Am I happy?** Based on these results, I am **[surprised/concerned]** because most simulations show **[gains/losses]** despite the positive expected value per flip.

## Question 5: Probability of Ending Up Above $1,000

**Answer:** The probability that your account balance will be greater than $1,000 at age 55 is approximately **`r round(prob_above_1000 * 100, 1)`%**.

This is calculated from our 100 simulations above - we simply count how many simulations ended up above $1,000 and divide by 100.

**Key Insight:** Even though each coin flip has a positive expected value, most people actually lose money over 30 years! This is the counterintuitive result of the investment game.

## Question 6: Modified Game Strategy Comparison

**Modified Game Rules:**
- You bet exactly 50% of your current balance on each flip
- Heads: You gain 25% of your current balance (1.25x multiplier)
- Tails: You lose 20% of your current balance (0.80x multiplier)

```{r}
#| label: modified-game-comparison
#| fig-cap: Comparison of original vs modified game strategies
#| echo: false

set.seed(456)
initial_balance <- 1000
n_years <- 30
n_sims <- 100

# Function for modified game
simulate_modified_game <- function(initial, years) {
  balance <- initial
  
  for (i in 1:years) {
    coin_flip <- rbinom(1, 1, 0.5)  # 1 = heads, 0 = tails
    
    if (coin_flip == 1) {
      balance <- balance * 1.25  # +25%
    } else {
      balance <- balance * 0.80  # -20%
    }
  }
  
  return(balance)
}

# Run 100 simulations for modified game
modified_final_balances <- replicate(n_sims, simulate_modified_game(initial_balance, n_years))

# Create comparison data frame
comparison_data <- data.frame(
  Game_Type = c(rep("Original Game", n_sims), rep("Modified Game", n_sims)),
  Final_Balance = c(final_balances, modified_final_balances)
)

# Create comparison histogram using ggplot2
library(ggplot2)
ggplot(comparison_data, aes(x = Final_Balance, fill = Game_Type)) +
  geom_histogram(alpha = 0.7, bins = 30, position = "identity") +
  geom_vline(xintercept = 10000, color = "red", linetype = "dashed", linewidth = 1.5) +
  scale_x_continuous(
    labels = scales::dollar_format(),
    trans = "log10",
    breaks = c(1, 10, 100, 1000, 10000, 100000)
  ) +
  labs(
    title = "Comparison of Final Balances: Original vs Modified Game",
    subtitle = "100 Simulations Each (Age 25 to 55)",
    x = "Final Balance at Age 55 ($)",
    y = "Frequency",
    caption = "Red line: $10,000 threshold"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12, color = "gray50"),
    legend.position = "bottom"
  ) +
  guides(fill = guide_legend(title = "Game Strategy"))

# Calculate probabilities
prob_original_above_10000 <- mean(final_balances > 10000)
prob_modified_above_10000 <- mean(modified_final_balances > 10000)

# Show comparison
cat("Strategy Comparison Results:\n")
cat("Original Game - Probability above $10,000: ", round(prob_original_above_10000 * 100, 1), "%\n")
cat("Modified Game - Probability above $10,000: ", round(prob_modified_above_10000 * 100, 1), "%\n")
cat("Original Game - Mean final balance: $", round(mean(final_balances), 2), "\n")
cat("Modified Game - Mean final balance: $", round(mean(modified_final_balances), 2), "\n")
```

**Answer:** 
- **Original Game:** Probability above $10,000 = **`r round(prob_original_above_10000 * 100, 1)`%**
- **Modified Game:** Probability above $10,000 = **`r round(prob_modified_above_10000 * 100, 1)`%**

**Comparison:** The probability in the modified game is **`r ifelse(prob_modified_above_10000 > prob_original_above_10000, "HIGHER", "LOWER")`** than the original game.

**Why?** The modified game is more conservative - smaller gains (+25% vs +50%) but also smaller losses (-20% vs -40%), which changes the risk profile.

# Conclusion: Key Insights 🎯

## What We Learned

**The Counterintuitive Truth:** Even though each coin flip has a positive expected value, most people actually lose money over 30 years!

### Key Insights:

1. **Expected Value ≠ Reality**: Just because the math says you should make money doesn't mean you will
2. **Volatility Kills Returns**: High volatility (big swings) reduces your chances of success over time
3. **Simulation Reveals Truth**: Running 100 simulations showed us what intuition couldn't predict

### Why This Happens:
- **Multiplicative Process**: Each loss compounds, making it harder to recover
- **Asymmetric Risk**: A 40% loss requires a 67% gain just to break even
- **Time Amplifies Problems**: Over 30 years, bad luck compounds dramatically

## The Big Picture

This challenge demonstrates why **simulation is so powerful** - it reveals insights that pure mathematics and intuition miss. The expected value said "you should make money," but the simulation said "most people lose money."

**Lesson:** When making important decisions, especially about money, don't just look at averages - understand the full range of possible outcomes through simulation.

---

*This analysis shows the power of computational methods in understanding complex systems where intuition alone can be misleading.*
